---
title: "Lab 1 EDLD 3"
author: "Vinita Vader & Raleigh Goodwin"
date: "4/2/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

```{r}
set.seed(123) # Set random number generation

m <- lm(mpg ~ hp + cyl + drat, mtcars)

l <- list(a = list(m = matrix(1:12, ncol = 3),
              	   v = 1:7),
          b = data.frame(student = 1:15,
                         score = rnorm(15, 100, 10)))
```

1. From the model results:

Extract the qr tolerance (tol).
```{r}
#str(m)
m$qr$tol
```

Extract the term labels from the model.
```{r}
#str(m)
attributes(m$terms)[3]
```

hint: Use str to inspect the list. This one is tricky and involves pulling an attribute from an attribute.

2. From the list

Use at least two different methods to extract m.
```{r}
#str(l)
l$a$m #method 1 - gives a df
l$a[1] #method 2 - this still looks like a list
l[[1]]$m #method 3 just because

```

Extract the third column of m. Maintain the dimensions (matrix structure).
```{r}
l$a$m[,3, drop=FALSE]
l$a$m[1:4,3, drop=FALSE]
```

Extract the score for student 7.
```{r}
l$b$score[7]
```

## Loops

3. Use the “airquality” dataset, which is part of base R. Split it by Month
```{r}
split(airquality, f = airquality$Month)#Method 1

airquality %>%   #Method 2
  group_split(Month)
```

4. Use a for loop to calculate the mean Ozone by month (remember to remove missing data from the calculation)
```{r}

data <- airquality %>%
  drop_na() %>% #drop na 
    group_split(Month)

# Oz_means <- rep(NA, length(levels(as.factor(data$Month)))) #5 levels of Month, hence vector of 5 open spots to be filled -- for some reason this was not working. although i know it's better to be specific to the variables since things can change length, for now, i'll just input length manually until i can troubleshoot

Oz_means <- rep(NA, 5) #5 levels of Month, hence vector of 5 open spots to be filled


for(i in seq_along(Oz_means)){
   Oz_means[i] <- data[[i]] %>% 
        summarise(mean(Ozone))
    
}

Oz_means
```

5. Replicate this calculation with lapply, sapply, and vapply
```{r}
lapply(data, function(x) mean(x$Ozone))

sapply(data, function(x) mean(x$Ozone))

vapply(data, function(x) mean(x$Ozone), FUN.VALUE = double(1))

```

6. Produce separate plots showing the relation between Ozone and Solar.R for each month (with a single loop).
```{r}
by_month <- split(airquality, airquality$Month) # needs to be a list

# str(by_month)

plots <- lapply(by_month, function(x) {
  ggplot(x, aes(Ozone, Solar.R)) +
    geom_point() +
    geom_smooth()
})

```

7. Use a for loop to save the plots to a folder on your computer
```{r}
dir.create(here::here("plots"))

filenames <- here::here("plots",
                        paste0("Month", names(by_month), ".png"))

filenames

for(i in seq_along(plots)) {
  ggsave(filenames[i],
         plots[[i]],
         device = "png",
         width = 9,
         height = 8)
}
```


